The Anatomy of a Temporal Summarization System

### 2015-05-29 dataset.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

Our experiments also make use of the TREC Temporal Summarization (TS) Track
 data from 2013 and 2014 \cite{aslam2013trec}. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event evaluation timeframe).  
All events occurred during the time span of the TREC KBA Stream Corpus.
For each event we create a stream of relevant documents by 
selecting only documents that contain the complete set 
of query words. 
%identifying the 
%article start/end within each web page in the corpus using a classifier
%tra
%those documents

Along with the metadata, NIST assessors constructed a set of ground truth nuggets for each event. 
Nuggets are brief and important text snippets that represent sub-events that should be conveyed
by an ideal update summary. % (see Figure~\ref{fig:nuggets} for examples).
In order
to accomplish this, for each event, assessors were provided with the
revision history of the Wikipedia page associated with the event.  
For example, 
the revision history for the Wikipedia page for `Hurricane Sandy' will 
contain text additions including those related to individual nuggets.  The assessment
task involves reviewing the Wikipedia revisions in the evaluation timeframe 
and marking the text additions capturing a new, unique nugget.  More detail
on this process can be found in the track description \cite{aslam2013trec}.

>>>>

Our experiments also make use of the TREC Temporal Summarization (TS) Track
 data from 2013 and 2014 \cite{aslam2013trec}.  This data includes 25 events 
and event metadata (e.g., a user search query for the event, the event type, 
and event evaluation time frame). All events occurred during the time span of 
the TREC KBA Stream Corpus. For each event we create a stream of relevant 
documents by selecting only documents that contain the complete set of query 
words.

Along with the metadata, NIST assessors constructed a set of ground truth 
nuggets for each event. Nuggets are brief and important text snippets that 
represent sub-events that should be conveyed by an ideal update summary. 
In order to accomplish this, for each event, assessors were provided with the
revision history of the Wikipedia page associated with the event. For example,
the revision history for the Wikipedia page for `Hurricane Sandy' will contain
text additions including those related to individual nuggets.  The assessment
task involves reviewing the Wikipedia revisions in the evaluation time frame 
and marking the text additions capturing a new, unique nugget.  More detail
on this process can be found in the track description \cite{aslam2013trec}.

### 2015-05-29 experiments.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. 
%Kernel length scale parameters  for each
%feature group are fit to maximize
%the model's log likelihood using ???.

@@

ROUGE measures the ngram overlap between a model summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
%See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE scores for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, 
%and so we can see how quickly
%each system is able to achieve its maximum performance.
and this evaluation reveals how quickly systems can cover  information in the model.

@@

%NIST developed a series metrics for evaluating update summarization systems
NIST developed metrics for evaluating update summarization systems
as part of the TREC TS track.
We present results on two of these metrics, the expected gain and 
comprehensiveness.

\textbf{Expected Gain } We treat the event's nuggets as unique units of 
information.
When a system adds an update to its summary, it is potentially adding some
of this nugget information. It would be instructive to know how much unique
and novel information each update is adding on average to the summary.
To that end, we define

\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
expected gain greater than 1.
%values greater than one can be thought of as a 
%compression ration, i.e. an expected gain of 10 indicates the average update
%contains 10 unique nuggets on average and that the summary is a 10~:~1 
%compression of the nugget information.
An expected gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure the completeness of an update 
summary. We define

\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has covered
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically; a
match exists if the semantic similarity of the update/nugget pair is above 
a threshold.
Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A manual inspection of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
%For example, the nugget 
%\begin{blockquote}
%watches and warnings in effect for much of the Mid-Atlantic states and 
%southern New England 
%\end{blockquote}
%\noindent
%matches with the update 
%\begin{blockquote}
%A tropical storm warning remains in effect for much of the coasts 
%of North and South Carolina\end{blockquote}
%\noindent
%with a similarity of .7. 
The average semantic similarity of manual matches performed 
by NIST assessors was much lower at approximately .25, increasing
our confidence in the automatic matches in the .5--1 range.

\subsection{Model Comparisons}
We refer to our complete model as \textbf{AP+Salience}.  We compare this model against several variants and baselines intended to measure the contribution of different components. All thresholds
for all runs are tuned on the development set.

\textbf{Affinity Propagation only (AP) } The purpose of this model
is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
clustering component, but without the salience information.
In this model, input sentences are \textit{apriori} equally likely to be exemplars;
%clustering algorithm without the salience predictions.
the salience values are uniformly set as the median value of the 
input similarity scores, as is commonly used in the AP literature \cite{frey2007clustering}.
After clustering a sentence batch, the exemplars are examined in order
of increasing time since event start and selected as updates if their
maximum similarity to the previous updates is less than $\lambda_{sim}$,
as in the novelty filtering stage of AP+Salience.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that AP+Salience is not just an improvement
over AP but centrality driven methods in general.
HAC was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
that the fusion of clustering and salience prediction improves over
predicting salience alone.
In this model we predict the salience of sentences as in step 1 for 
AP+Salience. %However, rather than clustering, we simply rank the sentences
%by salience.
We omit the clustering phase (step 2).
Updates are selected identically to step 3 of AP+Salience, proceeding 
in order of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.

%\fdcomment{consider removing AP+SalienceTR}
%\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
%to the
%AP+Salience model except at the novelty filter stage.
%Within the novelty filtering stage, 
%AP+Salience inspects exemplars in order of decreasing salience when computing
%similarity to previous updates. Because the order in which we add updates
%affects future similarity calculations, we evaluate an alternate ordering 
%scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
%time since event start. The purpose of this model is to give preference to 
%more recent updates, since timelyness is one of the desirable traits of our update
%summaries.
%

>>>>

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream.

@@

ROUGE measures the ngram overlap between a model summary and an automatically 
generated system summary. Model summaries for each event were constructed by 
concatenating the event's nuggets. Generally, ROUGE evaluation assumes both 
model and system summaries are of a bounded length. Since our systems are 
summarizing events over a span of two weeks time, the total length of our 
system output is much longer than the model. To address this, for each 
system/event pair, we sample with replacement 1000 random summaries of length 
less than or equal to the model summary (truncating the last sentence when 
neccessary). The final ROUGE scores for the system are the average scores 
from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. The model 
summaries in this case are retrospective, and this evaluation reveals how 
quickly systems can cover  information in the model.

@@

NIST developed metrics for evaluating update summarization systems as part of 
the TREC TS track. We present results on two of these metrics, the expected 
gain and comprehensiveness.

\textbf{Expected Gain } We treat the event's nuggets as unique units of 
information. When a system adds an update to its summary, it is potentially 
adding some of this nugget information. It would be instructive to know how 
much unique and novel information each update is adding on average to the 
summary. To that end, we define
\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}
\] 
where $S$ is the set of system updates, $S_n$ is the set of nuggets contained 
in $S$, and $|\cdot|$ is the number of elements in the set. To compute the 
set $S_n$ we match each system update to 0 or more nuggets, where an update 
matches a nugget if their semantic similarity is above a threshold. $S_n$ 
results from the unique set of nuggets matched. Because an update can map to 
more than one nugget, it is possible to receive an expected gain greater 
than 1. An expected gain of 1 would indicate that every sentence was both 
relevant and contained a unique piece of information.

\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure 
the completeness of an update summary. We define
\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]
where $N$ is the set of event nuggets. A comprehensiveness of 1 indicates that
the summary has covered all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically; a match exists if the 
semantic similarity of the update/nugget pair is above a threshold. 
Determining an optimal threshold to count matches is difficult so we evaluate 
at threshold values ranging from .5 to 1, where values closer to 1 are more 
conservative estimates of performance. A manual inspection of matches suggests
that semantic similarity values around .7 produce reasonable matches.
The average semantic similarity of manual matches performed by NIST assessors 
was much lower at approximately .25, increasing our confidence in the 
automatic matches in the .5--1 range.

\subsection{Model Comparisons}

We refer to our complete model as \textsc{AP+Salience}.  We compare this model
against several variants and baselines intended to measure the contribution of
different components. All thresholds for all runs are tuned on the development
set.

\textbf{Affinity Propagation only (\textsc{AP}) } The purpose of this model 
is to 
directly measure the effect of integrating salience and clustering by 
providing a baseline that uses the identical clustering component, but without
the salience information. In this model, input sentences are \textit{apriori}
equally likely to be exemplars; the salience values are uniformly set as the 
median value of the  input similarity scores, as is commonly used in the 
\textsc{AP} 
literature \cite{frey2007clustering}. After clustering a sentence batch, the 
exemplars are examined in order of increasing time since event start and 
selected as updates if their maximum similarity to the previous updates is 
less than $\lambda_{sim}$, as in the novelty filtering stage of 
\textsc{AP+Salience}.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another 
clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that \textsc{AP+Salience} is not just an 
improvement
over \textsc{AP} but centrality driven methods in general. \textsc{HAC} was 
chosen over other 
clustering approaches because the number of clusters is not an explicit 
hyper-parameter. To produce flat clusters from the hierarchical clustering, we
flatten the \textsc{HAC} dendrogram using the cophenetic distance criteria, 
i.e. 
observations in each flat cluster have no greater a cophenetic distance than a
threshold. Cluster centers are determined to be the sentence with highest 
cosine similariy to the flat cluster mean. Cluster centers are examined in 
time order and are added to the summary if their similarity to previous 
updates is below a similarity threshold $\lambda_{sim}$ as is done in the 
\textsc{AP} model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
that the fusion of clustering and salience prediction improves over
predicting salience alone. In this model we predict the salience of sentences 
as in step 1 for \textsc{AP+Salience}. We omit the clustering phase (step 2).
Updates are selected identically to step 3 of \textsc{AP+Salience}, 
proceeding in order
of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.

### 2015-05-29 introduction.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

\label{sec:introduction}
During crises, information is critical for first responders,
 crisis management organizations, and those caught in the event.  
When the event is significant, as in the case of Hurricane
Sandy, the amount of content produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. 
Crisis informatics \cite{palen2010vision} is dedicated to finding methods for sharing the
right information in a timely fashion during such an event.
Research in this field has focused on
human-in-the-loop approaches ranging from on the ground information 
gathering to crowdsourced reporting and disaster management \cite{starbird2013working}.

Multi-document summarization has the potential to assist the crisis 
informatics community. 
Automatic summarization could deliver
relevant and salient information at regular intervals, 
even when human volunteers are unable to. 
Perhaps more importantly it could help filter out
unnecessary and irrelevant detail when the volume of incoming information
is large. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
these experiments were not developed to handle streaming data from a
 heterogeneous environment at web scale. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering or relevance alone.  
Our experiments suggest that this
is because our system is better able to adapt to
dynamic changes in input volume that adversely 
affect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disaster, this approach is generally applicable to many update summarization tasks.
% where queries can be categorized by the 
%expected similarity of their ideal summaries.
%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
%expected similarity of their ideal summaries.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

Our approach achieves a statistically significant improvement in 
 ROUGE scores compared to multiple baselines.
Additionally, we introduce novel methods for estimating the average information
gain each update provides and how completely the update summary covers 
the event it is tracking; our system's updates contain more relevant
information on average than the competeing baselines.

@@

%Our main contributions are three fold. First, we present a straightforward 
%framework for 
%combining prior beliefs with a clustering algorithm. Second, we demonstrate
%the empirical effectiveness of this approach on the update summarization
%task. Finally, we present a novel feature represention for sentence salience
%in the crisis/disaster domain.

>>>>

\label{sec:introduction}

During crises, information is critical for first responders, crisis management
organizations, and those caught in the event. When the event is significant, 
as in the case of Hurricane Sandy, the amount of content produced by 
traditional news outlets, government agencies, relief organizations, and 
social media can vastly overwhelm those trying to monitor the situation. 
Crisis informatics \cite{palen2010vision} is dedicated to finding methods for 
sharing the right information in a timely fashion during such an event.
Research in this field has focused on human-in-the-loop approaches ranging 
from on the ground information gathering to crowdsourced reporting and 
disaster management \cite{starbird2013working}.

Multi-document summarization has the potential to assist the crisis 
informatics community. Automatic summarization could deliver relevant and 
salient information at regular intervals, even when human volunteers are 
unable to. Perhaps more importantly it could help filter out unnecessary and 
irrelevant detail when the volume of incoming information is large. While 
methods for identifying, tracking, and summarizing events from text based 
input have been explored extensively
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
these experiments were not developed to handle streaming data from a
heterogeneous environment at web scale. These methods also rely heavily on 
redundancy which is suboptimal for time sensitive domains where there is a 
high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. We demonstrate that 
combining salience with clustering produces more relevant summaries compared 
to baselines using clustering or relevance alone.  Our experiments suggest 
that this is because our system is better able to adapt to dynamic changes in 
input volume that adversely affect methods that use redundancy as a proxy for 
salience.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disasters, this approach is 
generally applicable to many update summarization tasks.

Our approach achieves a statistically significant improvement in ROUGE scores 
compared to multiple baselines. Additionally, we introduce novel methods for 
estimating the average information gain each update provides and how 
completely the update summary covers the event it is tracking; our system's 
updates contain more relevant information on average than the competing 
baselines.

@@
